### 2022-07-31 manuscript/appendix_rtt.tex
add reference, 19B++ clade

\section*{Divergence summaries}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/19A_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 19A.}
    \label{fig:19A_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/19B_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 19B.}
    \label{fig:19B_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/19B+_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 19B+.}
    This figure contains sequences in clades 19 A and B rooted on clade 19B.
    \label{fig:19B+_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/20A_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 20A.}
    \label{fig:20A_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/20B_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 20B.}
    \label{fig:20B_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/20C_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 20C.}
    \label{fig:20C_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/20A+_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 20A+.}
    This figure contains sequences in clades 20A,B,C,D rooted on clade 20A.
    \label{fig:20A+_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/20E_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 20E.}
    \label{fig:20E_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/20H_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 20H (Beta).}
    \label{fig:20H_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/20I_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 20I (Alpha).}
    \label{fig:20I_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/20J_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 20J (Gamma).}
    \label{fig:20J_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/21D_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 21D (Eta).}
    \label{fig:21D_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/21G_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 21G (Lambda).}
    \label{fig:21G_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/21H_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 21H (Mu).}
    \label{fig:21H_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/21I_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 21I (Delta).}
    \label{fig:21I_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/21J_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 21J (Delta).}
    \label{fig:21J_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/21K_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 21K (Omicron).}
    \label{fig:21K_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/21L_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 21L (Omicron).}
    \label{fig:21L_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/22A_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 22A (Omicron).}
    \label{fig:22A_divergence}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/rtt/22B_rtt.png}
    \caption{{\bf Divergence increases linearly with time in clade 22B (Omicron).}
    \label{fig:22B_divergence}}
\end{figure*}