### 2022-08-02 manuscript/appendix.tex
start fitness section

\section*{Supplementary figures}

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/mutation_distribution.pdf}
    \caption{{\bf The relative rate of different mutations in SARS-CoV-2.}
    These rates are measured from rare low frequency mutations probably subject to little purifying selection.
    \label{fig:mutation_distribution}}
\end{figure*}